Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean throughout the United States, and from Mexico, France, and Australia

### 2017-06-27 doc/manuscript/manuscript.Rmd
minor edits

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in Mexico, France, Australia, and 11 states of the United States of America (Table 1).
The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table 1).
The bean nurseries were 5 m x 10 m in size and maintained without applying fungicides to observe natural incidence of white mold disease.
The nursery plots were incorporated with a basal dressing of N: P: K = 1: 3: 2 and side dressing of 0:3:2 during the season.
Although the nursery locations were the same over sampling years, sampling fields within a location varied for sampling years.
Sampling was carried out on a 'W' or 'V' shape transect, or randomly.
Sampling year and number of isolates collected were necessarily influenced by incidence and severity of white mold disease.

@@

Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig. 3Z): A costal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Supplementary Figure S1Z).
The international community appears to be driven by MLG 4, which is shared between all three populations and has a Psex value of 2.87 x 10-5, in contrast to the abundant MLG 25, which has a Psex value of 0.0168.

@@

The dapc for hosts was limited to six most common cultivars: G122 (N = 55), Beryl (N = 62), Bunsi (N = 47), Yellow (N = 26), Pinto (N = 15), and Merlot (N = 11).
According to Fig. 8B, cultivar Merlot isolates group distant from others, but were not significantly different (P =0.325).  Interestingly, *S. sclerotiorum* populations from cultivars 'Beryl' and 'G122', which are known to have different level of white mold susceptibility, showed no difference in population genetic data.

@@

Contrary to above results Aldrich-Wolfe et al. (2015) who did a survey on *S. sclerotiorum* from different host crops in the North Central states of the United States found a single microsatellite haplotype to be prevalent throughout the region.
A study in the UK that examined this pathogen from 12 populations in different crop plants and weed species *Ranunculus acris* (meadow buttercup) also reported one microsatellite haplotype was widely distributed across hosts and areas (Clarkson et al., 2013).
Reason for our isolates from dry bean not to show a widespread MLG may be due to extensive sexual outcrossing and movement sclerotia by human activities.
Though conditions of outcrossing for *S. sclerotiorum* is not well studied, it is believed weather and agricultural conditions play a role in production and release of ascospores of this pathogen (Huang, 1991; Sexton et al., 2006).
One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% (Clarkson et al., 2003).
The same study reported increased ascospores discharge at 20-25° C compared to 15° C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
Our samples were collected from 2003 to 2012 and except for the dry bean nurseries, came from different fields.
Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
When conditions are favorable for ascospore production, release, and survival, coinfection of a single site on a plant by multiple ascospores is possible (Sexton et al., 2006).
This may eventually result in facilitating outcrossing and recombination.

@@

We performed DAPC without assigning isolates to different populations based on sampling information to see the how isolates fall into clusters according to their genetic relatedness.
We decided the most appropriate number as six since the DAPC results had 87% (N = 13/15) of Mexican isolates in a single cluster (Fig. 9 cluster 3).
When number of clusters was assigned less than that MX isolates did not group separately while higher number of cluster assignment resulted in breaking down of other clusters in Fig. 9 into smaller groups.
DAPC analysis without population assignments revealed that in general geography is not a barrier for gene flow since isolates of different geographies from USA, Australia and France clustered together indicating genetic relatedness and possible sexual recombination within each region.
For instance France isolates were present in cluster 1, 2, 4 and 5.
As mentioned above only exception was the Mexican isolates.
Njambere et al. (2014) studied S. trifoliorum, a close relative *S. sclerotiorum* and reported four cryptic genetic populations by Bayesian clustering analysis of microsatellite loci.
Those clusters had no correlation to geographic locations of samples.
Similarly, a cluster analysis of microsatellite data of *S. sclerotiorum* collected from 12 geographically distant host populations in  England and Wales resulted only two main clusters pooling isolates from different populations together (Clarkson et al., 2013).
However, some other studies found *S. sclerotiorum* isolates from distant locations did differ significantly when molecularly characterized (Attanayake et al., 2013; Cubeta et al., 1997; Ekins, 1999; Malvarez et al., 2007).

@@

*Sclerotinia sclerotiorum* isolates from dry bean represented high microsatellite diversity and lack of association MCGs and microsatellite MLGs.
Analysis of molecular variance of populations based on region and sampling year showed low level of differentiation with majority of variance detected within populations.
These features are indicative of sexual outcrossing and perhaps human mediated migration of *S. sclerotiorum* in dry bean.
However, we also detected clones or identical microsatellite MLGs within and among populations indicating asexual reproduction.
Further, linkage disequilibrium and index of association calculations suggested *S. sclerotiorum* populations are largely clonal.
Although, there was low impact of region and year on population differentiation, we did not detect host cultivar as a barrier to gene flow of the pathogen.
Our study did not find a widespread genotype among sampled regions from dry beans but some Australian and France MLGs were found in the USA isolates.
The analysis of populations from same field over different years for linkage disequilibrium did not provide evidence for sexual outcrossing.
However, analysis of molecular variance did show populations were significantly different between some years (eg.
MI 2003 and 2005 populations; WA 2008 compared to both 2003 and 2005).
These results indicate some level of heterothalic recombination is taking place in these fields in addition to natural mutations.
An understanding of sexual behavior and ability of form new genotypes by *S. sclerotiorum* in dry bean cropping systems is useful for formulating disease management strategies.

>>>>

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in Mexico, France, Australia, and 11 states of the United States of America (Table 1).
The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table 1).
The bean nurseries were 5 m x 10 m in size and maintained without applying fungicides to observe natural incidence of white mold disease.
The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the season.
Although the nursery locations were the same over sampling years, sampling fields within a location varied for sampling years.
Sampling was carried out on a 'W' or 'V' shape transect, or randomly.
Sampling year and number of isolates collected were necessarily influenced by incidence and severity of white mold disease.

@@

Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig. 3Z): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Supplementary Figure S1Z).
The international community appears to be driven by MLG 4, which is shared between all three populations and has a Psex value of 2.87 x 10-5, in contrast to the abundant MLG 25, which has a Psex value of 0.0168.

@@

The DAPC for hosts was limited to six most common cultivars: G122 (N = 55), Beryl (N = 62), Bunsi (N = 47), Yellow (N = 26), Pinto (N = 15), and Merlot (N = 11).
According to Fig. 8B, cultivar Merlot isolates group distant from others, but were not significantly different (P =0.325).  Interestingly, *S. sclerotiorum* populations from cultivars 'Beryl' and 'G122', which are known to have different level of white mold susceptibility, showed no difference in population genetic data.

@@

Contrary to above results Aldrich-Wolfe et al. (2015) who did a survey on *S. sclerotiorum* from different host crops in the North Central states of the United States found a single microsatellite haplotype to be prevalent throughout the region.
A study in the UK that examined this pathogen from 12 populations in different crop plants and weed species *Ranunculus acris* (meadow buttercup) also reported one microsatellite haplotype was widely distributed across hosts and areas (Clarkson et al., 2013).
Reason for our isolates from dry bean not to show a widespread MLG may be due to extensive sexual outcrossing and movement sclerotia by human activities.
Though conditions of outcrossing for *S. sclerotiorum* is not well studied, it is believed weather and agricultural conditions play a role in production and release of ascospores of this pathogen (Huang, 1991; Sexton et al., 2006).
One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% (Clarkson et al., 2003).
The same study reported increased ascospores discharge at 20-25° C compared to 15° C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
Our samples were collected from 2003 to 2012 and except for the dry bean nurseries, came from different fields.
Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
When conditions are favorable for ascospore production, release, and survival, co-infection of a single site on a plant by multiple ascospores is possible (Sexton et al., 2006).
This may eventually result in facilitating outcrossing and recombination.

@@

We performed DAPC without assigning isolates to different populations based on sampling information to see the how isolates fall into clusters according to their genetic relatedness.
We decided the most appropriate number as six since the DAPC results had 87% (N = 13/15) of Mexican isolates in a single cluster (Fig. 9 cluster 3).
When number of clusters was assigned less than that MX isolates did not group separately while higher number of cluster assignment resulted in breaking down of other clusters in Fig. 9 into smaller groups.
DAPC analysis without population assignments revealed that in general geography is not a barrier for gene flow since isolates of different geographies from USA, Australia and France clustered together indicating genetic relatedness and possible sexual recombination within each region.
For instance France isolates were present in cluster 1, 2, 4 and 5.
As mentioned above only exception was the Mexican isolates.
Njambere et al. (2014) studied *S. trifoliorum*, a close relative *S. sclerotiorum* and reported four cryptic genetic populations by Bayesian clustering analysis of microsatellite loci.
Those clusters had no correlation to geographic locations of samples.
Similarly, a cluster analysis of microsatellite data of *S. sclerotiorum* collected from 12 geographically distant host populations in  England and Wales resulted only two main clusters pooling isolates from different populations together (Clarkson et al., 2013).
However, some other studies found *S. sclerotiorum* isolates from distant locations did differ significantly when molecularly characterized (Attanayake et al., 2013; Cubeta et al., 1997; Ekins, 1999; Malvarez et al., 2007).

@@

*Sclerotinia sclerotiorum* isolates from dry bean represented high microsatellite diversity and lack of association MCGs and microsatellite MLGs.
Analysis of molecular variance of populations based on region and sampling year showed low level of differentiation with majority of variance detected within populations.
These features are indicative of sexual outcrossing and perhaps human mediated migration of *S. sclerotiorum* in dry bean.
However, we also detected clones or identical microsatellite MLGs within and among populations indicating asexual reproduction.
Further, linkage disequilibrium and index of association calculations suggested *S. sclerotiorum* populations are largely clonal.
Although, there was low impact of region and year on population differentiation, we did not detect host cultivar as a barrier to gene flow of the pathogen.
Our study did not find a widespread genotype among sampled regions from dry beans but some Australian and France MLGs were found in the USA isolates.
The analysis of populations from same field over different years for linkage disequilibrium did not provide evidence for sexual outcrossing.
However, analysis of molecular variance did show populations were significantly different between some years (eg.
MI 2003 and 2005 populations; WA 2008 compared to both 2003 and 2005).
These results indicate some level of heterothallic recombination is taking place in these fields in addition to natural mutations.
An understanding of sexual behavior and ability of form new genotypes by *S. sclerotiorum* in dry bean cropping systems is useful for formulating disease management strategies.
